---
title: "Assignment 1"
author: "Group 1"
date: "2025-11-29"
output:
  html_document:
    df_print: paged
---

---
title: "Students Performance Data Analysis"
author: "Group 1"
date: "`r format(Sys.Date())`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
fontsize: 11pt
geometry: margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
set.seed(123)
```

## Load Packages

```{r libraries}
library(tidyverse)
library(janitor)
library(psych)
library(rsample)
```

## Import Data

```{r import-data}
students <- read_csv("C:/Users/Bharat/Desktop/Proj/StudentsPerformance.csv") %>%
  clean_names()

dim(students)
names(students)
head(students, 10)
```

## Structure and Variables

```{r structure}
str(students)
summary(students)
```

## Filter Rows (Example: students with math score > 80)

```{r filter-rows}
high_math <- students %>%
  filter(math_score > 80)

head(high_math, 10)
```

## Remove Missing Values

```{r remove-missing}
students_no_na <- students %>% drop_na()
sum(!complete.cases(students_no_na))
```

## Remove Duplicates

```{r remove-duplicates}
students_dedup <- students_no_na %>% distinct()
nrow(students_no_na) - nrow(students_dedup)
```

## Reorder Rows (Descending by reading score)

```{r reorder-rows}
students_ordered <- students_dedup %>% arrange(desc(reading_score))
head(students_ordered, 10)
```

## Rename Columns (for easier use)

```{r rename-columns}
students_renamed <- students_dedup %>%
  rename(Gender = gender,
         Ethnicity = race_ethnicity,
         Parental_Education = parental_level_of_education,
         Lunch = lunch,
         Test_Prep = test_preparation_course,
         Math = math_score,
         Reading = reading_score,
         Writing = writing_score)

names(students_renamed)
```

## Add New Variables

```{r add-vars}
students_augmented <- students_renamed %>%
  mutate(Average_Score = (Math + Reading + Writing)/3,
         Passed_Math = ifelse(Math >= 50, "Yes", "No"))

head(students_augmented, 10)
```

## Training and Testing Split

```{r train-test}
split_obj <- initial_split(students_augmented, prop = 0.7)
train_df <- training(split_obj)
test_df  <- testing(split_obj)

nrow(train_df); nrow(test_df)
```

## Summary Statistics

```{r summary-stats}
students_augmented %>%
  group_by(Gender) %>%
  summarize(
    count = n(),
    mean_math = mean(Math),
    mean_reading = mean(Reading),
    mean_writing = mean(Writing)
  )
```

## Mean, Median, Mode, Range (Math Scores)

```{r mean-median-mode-range}
x <- students_augmented$Math

mean(x)
median(x)

mode_stat <- function(v) {
  v <- v[!is.na(v)]
  u <- unique(v)
  u[which.max(tabulate(match(v, u)))]
}
mode_stat(x)

range(x)
```

## Scatter Plot (Math vs Reading)

```{r scatter-plot, fig.height=4, fig.width=6}
ggplot(students_augmented, aes(x = Math, y = Reading, color = Gender)) +
  geom_point(alpha = 0.8) +
  labs(title = "Scatter: Math vs Reading Scores") +
  theme_minimal()
```

## Bar Plot (Average Score by Lunch Type)

```{r bar-plot, fig.height=4, fig.width=6}
students_augmented %>%
  group_by(Lunch) %>%
  summarize(mean_avg = mean(Average_Score)) %>%
  ggplot(aes(x = Lunch, y = mean_avg, fill = Lunch)) +
  geom_col() +
  labs(title = "Bar: Average Score by Lunch Type") +
  theme_minimal()
```

## Pearson Correlation (Math vs Writing)

```{r correlation}
cor(students_augmented$Math, students_augmented$Writing, method = "pearson")
```

---

## References

Dataset source: Kaggle — Students Performance Dataset  
GitHub repository: [https://github.com/bharatchopra-tech/Studentdata-analysisGroup1](https://github.com/bharatchopra-tech/Studentdata-analysisGroup1)