---
title: "Iris Data Analysis Project"
author: "Group 1"
date: "2025-11-29"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
set.seed(123)
```

## Load Packages

```{r libraries}
library(tidyverse)
library(janitor)
library(psych)
library(rsample)
```

## Import Data

```{r import-data}
iris_data <- read_csv("iris.csv") %>%
  clean_names()

dim(iris_data)
names(iris_data)
head(iris_data, 15)
```

## User-defined Function

```{r normalize-function}
normalize <- function(x) {
  (x - min(x, na.rm = TRUE)) / (max(x, na.rm = TRUE) - min(x, na.rm = TRUE))
}

iris_data <- iris_data %>%
  mutate(sepal_length_norm = normalize(sepal_length_cm))
```

## Filtering Rows

```{r filter-rows}
iris_filtered <- iris_data %>%
  filter(petal_length_cm > 4)

head(iris_filtered, 10)
```

## Remove Missing Values

```{r remove-missing}
iris_no_na <- iris_data %>% drop_na()
sum(!complete.cases(iris_no_na))
```

## Remove Duplicates

```{r remove-duplicates}
iris_dedup <- iris_no_na %>% distinct()
nrow(iris_no_na) - nrow(iris_dedup)
```

## Reorder Rows

```{r reorder-rows}
iris_ordered <- iris_dedup %>% arrange(desc(petal_width_cm), desc(sepal_length_cm))
head(iris_ordered, 10)
```

## Rename Columns

```{r rename-columns}
iris_renamed <- iris_dedup %>%
  rename(Sepal_Length = sepal_length_cm,
         Sepal_Width  = sepal_width_cm,
         Petal_Length = petal_length_cm,
         Petal_Width  = petal_width_cm)

names(iris_renamed)
```

## Add New Variables

```{r add-vars}
iris_augmented <- iris_renamed %>%
  mutate(double_petal_length = Petal_Length * 2,
         sepal_area_approx = Sepal_Length * Sepal_Width)

head(iris_augmented, 10)
```

## Training and Testing Split

```{r train-test}
split_obj <- initial_split(iris_augmented, prop = 0.7)
train_df <- training(split_obj)
test_df  <- testing(split_obj)

nrow(train_df); nrow(test_df)
```

## Summary Statistics

```{r summary-stats}
summary(iris_augmented)
```

## Mean, Median, Mode, Range

```{r mean-median-mode-range}
x <- iris_augmented$Sepal_Length

mean(x)
median(x)

mode_stat <- function(v) {
  v <- v[!is.na(v)]
  u <- unique(v)
  u[which.max(tabulate(match(v, u)))]
}
mode_stat(x)

range(x)
```

## Scatter Plot

```{r scatter-plot, fig.height=4, fig.width=6}
ggplot(iris_augmented, aes(x = Sepal_Length, y = Petal_Length)) +
  geom_point(alpha = 0.8, color = "steelblue") +
  labs(title = "Scatter: Sepal Length vs Petal Length") +
  theme_minimal()
```

## Bar Plot (Example: Frequency of Rounded Sepal Lengths)

```{r bar-plot, fig.height=4, fig.width=6}
iris_augmented %>%
  mutate(sepal_length_round = round(Sepal_Length)) %>%
  count(sepal_length_round) %>%
  ggplot(aes(x = sepal_length_round, y = n)) +
  geom_col(fill = "darkorange") +
  labs(title = "Bar: Frequency of Rounded Sepal Lengths",
       x = "Sepal Length (rounded)", y = "Count") +
  theme_minimal()
```

## Pearson Correlation

```{r correlation}
cor(iris_augmented$Sepal_Length, iris_augmented$Petal_Length, method = "pearson")
```

---

## References

Dataset source: Kaggle — Iris Dataset (`iris.csv`)  
GitHub repository: [https://github.com/bharatchopra-tech/Netflixldata-analysisGroup1](https://github.com/bharatchopra-tech/Netflixldata-analysisGroup1)